---
title: "BRM Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in raw data, cleaning

```{r}
pupil_measures <- readxl::read_excel("../data/Engagement_Data.xlsx",
                                 sheet = "Event_Statistics") %>%
  filter(Category == "Fixation") %>% 
  select(Trial, video, Export.End.Trial.Time..ms., Participant, Tracking.Ratio...., Category, Eye.L.R, Event.Start.Trial.Time..ms., Event.End.Trial.Time..ms., Event.Duration..ms., Fixation.Average.Pupil.Diameter..mm., type) %>%
  rename(trial = Trial,
         length_trial = Export.End.Trial.Time..ms.,
         participant = Participant,
         track_ratio = Tracking.Ratio....,
         eye = Eye.L.R,
         measure_start = Event.Start.Trial.Time..ms.,
         measure_end = Event.End.Trial.Time..ms.,
         duration = Event.Duration..ms.,
         pupil_mm = Fixation.Average.Pupil.Diameter..mm.) %>% 
  filter(length_trial > 500000) 
```

Reading in baseline data

```{r}
baseline <- read_csv("../data/baseline_data.csv")
```

Plotting raw pupil measurements over time

```{r}
pupil_measures %>% 
  ggplot(aes(x=measure_start, y = pupil_mm)) +
  geom_point(aes(col = participant))
```

Change from baseline

```{r}
pupil_measures %>% 
  select(participant, measure_start, pupil_mm) %>% 
  left_join(baseline %>% 
              select(participant, pupil_dia_mm_mean), by = c("participant")) %>% 
  mutate(pupil_change = pupil_mm - pupil_dia_mm_mean) %>% 
  ggplot(aes(x=measure_start, y = pupil_change)) +
  geom_point(aes(col = participant))
```

Clean pupil data based on:https://link.springer.com/article/10.3758/s13428-018-1075-y

Remove values outside of feasible range: 1.5-9 mm


```{r}
pupil_measures <- pupil_measures %>% 
  filter(!(pupil_mm < 1.5 | pupil_mm > 9))
```

Filtering outliers based on dilation speed

1) Estimate dilation speed

```{r}
pupil_measures <- pupil_measures %>% 
  arrange(participant, eye, measure_start) %>%
  group_by(participant, eye) %>% 
   mutate(d_lag = abs((pupil_mm - lag(pupil_mm)) / (measure_start - lag(measure_start))),
         d_lead = abs((pupil_mm - lead(pupil_mm)) / (measure_start - lead(measure_start)))) %>% 
  mutate(d_speed = pmax(d_lag, d_lead, na.rm = TRUE)) %>% 
  select(-c(d_lag, d_lead)) %>% 
  ungroup()
```

2) Assessing distribution of deviations from median

```{r}
pupil_measures %>% 
  mutate(d_deviation = abs(d_speed - median(d_speed, na.rm = TRUE))) %>% 
  ggplot() +
  geom_histogram(aes(x = d_deviation))
```

3) Estimating mean absolute deviation across participants

```{r}
pupil_measures %>% 
  mutate(d_deviation = abs(d_speed - median(d_speed, na.rm = TRUE))) %>% 
  summarise(MAD = median(d_deviation, na.rm = TRUE),
            threshold = median(d_speed, na.rm = TRUE) + 10*MAD)
```

4) Filtering observtions above threshold

```{r}
pupil_measures <- pupil_measures %>% 
  filter(!(d_speed > 0.002303483))
```


Functional Mixed Effections modeling