---
title: "BRM Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in raw data, cleaning

```{r}
pupil_measures <- read_csv("../data/pupil_data.csv") %>%
  filter(event == "Fixation") %>%
  filter(trial_length_ms > 500000) %>% 
  mutate(pupil_diameter = as.numeric(pupil_diameter))
```

Reading in baseline data

```{r}
baseline <- read_csv("../data/baseline_data.csv")
```

Plotting raw pupil measurements over time

```{r}
pupil_measures %>% 
  ggplot(aes(x=event_start_ms, y = pupil_diameter)) +
  geom_point(aes(col = participant))
```

```{r}
pupil_measures %>% 
  ggplot(aes(x=event_start_ms, y = pupil_diameter)) +
  geom_line(aes(col = participant))
```

Change from baseline

```{r}
pupil_measures %>% 
  select(participant, event_start_ms, pupil_diameter) %>% 
  left_join(baseline %>% 
              select(participant, pupil_dia_mm_mean), by = c("participant")) %>% 
  mutate(pupil_change = pupil_diameter - pupil_dia_mm_mean) %>% 
  ggplot(aes(x=event_start_ms, y = pupil_change)) +
  geom_point(aes(col = participant))
```

Clean pupil data based on:https://link.springer.com/article/10.3758/s13428-018-1075-y

Remove values outside of feasible range: 1.5-9 mm


```{r}
pupil_measures <- pupil_measures %>% 
  filter(!(pupil_diameter < 1.5 | pupil_diameter > 9))
```

Filtering outliers based on dilation speed

1) Estimate dilation speed

```{r}
pupil_measures <- pupil_measures %>% 
  arrange(participant, eye, event_start_ms) %>%
  group_by(participant, eye) %>% 
   mutate(d_lag = abs((pupil_diameter - lag(pupil_diameter)) / (event_start_ms - lag(event_start_ms))),
         d_lead = abs((pupil_diameter - lead(pupil_diameter)) / (event_start_ms - lead(event_start_ms)))) %>% 
  mutate(d_speed = pmax(d_lag, d_lead, na.rm = TRUE)) %>% 
  select(-c(d_lag, d_lead)) %>% 
  ungroup()
```

2) Assessing distribution of deviations from median

```{r}
pupil_measures %>% 
  mutate(d_deviation = abs(d_speed - median(d_speed, na.rm = TRUE))) %>% 
  ggplot() +
  geom_histogram(aes(x = d_deviation))
```

3) Estimating mean absolute deviation across participants

```{r}
pupil_measures %>% 
  mutate(d_deviation = abs(d_speed - median(d_speed, na.rm = TRUE))) %>% 
  summarise(MAD = median(d_deviation, na.rm = TRUE),
            threshold = median(d_speed, na.rm = TRUE) + 10*MAD)
```

4) Filtering observtions above threshold

```{r}
pupil_measures <- pupil_measures %>% 
  filter(!(d_speed > 0.002303483))
```


Functional Mixed Effections Modelling

Check for packages actually used

```{r}
library(lme4)
library(nlme)
library(splines)
library(lmerTest)
library(MuMIn)
```

Minimum trial length: truncating videos here

```{r}
trial_min <- pupil_measures %>% 
  summarise(trial_min = min(trial_length_ms)) %>% 
  .$trial_min
```

Resample 1 second intervals

```{r}
test_measures <- pupil_measures %>% 
  mutate(sec_1 = event_start_ms %/% 1000) %>% 
  filter(event_start_ms < trial_min) %>% 
  group_by(participant, sec_1) %>% 
  summarise(pupil_diameter = mean(pupil_diameter, na.rm = TRUE))
```

Function to plot resulting model

```{r}

plot_fda <- function(model){
  
g <- test_measures %>%
  ggplot(aes(x = sec_5, y = pupil_diameter)) +
  geom_point(size = 1, col = "darkgrey") +
  geom_line(aes(y = predict(model), group = participant, col = "Subject"), size = 0.5) +
  geom_line(data = test_measures %>% distinct(sec_5), aes(y = predict(model, re.form = NA, newdata = test_measures %>% distinct(sec_5)), col="Population"), size = 1.5) +
  scale_colour_manual(values = c("Subject" = "darkgrey", "Population" = "red")) +
  xlab("1-second interval") +
  ylab("Pupil Size (mm)") +
  #scale_x_continuous(labels = c("0:00", "1:00", "2:00", "3:00", "4:00", "5:00", "6:00", "7:00", "8:00", "9:00"), breaks = c(0, seq(3,27,3))) +
  scale_y_continuous(breaks = seq(0,5,0.5), limits = c(1.5,5.5)) +
  theme(legend.title = element_blank()) + 
    theme(axis.text = element_text(size = 6)) + 
    theme(axis.title = element_text(size = 8)) +
    theme(legend.direction = "horizontal") +
    theme(legend.position = "bottom") +
    theme(legend.margin = margin(t = 0, b = 0, unit='cm')) +
    theme(axis.title = element_text(size = 10)) +
    theme(axis.text = element_text(size = 8))

return(g)

}
```

### AIC as number of knots is increased

Considering 1 to 14 knots placed at quantiles of time intervals.

```{r, warning=FALSE, message=FALSE}

AIC_s <- rep(0,14)
BIC_s <- rep(0,14)

for (i in 2:15){
  
  PS_model <- lmer(data = test_measures,
                       pupil_diameter ~ bs(sec_1, df = i, degree = 1) + (bs(sec_1, df = i, degree = 1) | participant))

  AIC_s[i-1] <- AICc(PS_model)
  BIC_s[i-1] <- BIC(PS_model)
}
```

Plotting AICc and BIC vs. Number of Knots

```{r}
ggplot(data = data_frame(knots = 1:14, AIC = AIC_s, BIC = BIC_s)) +
  geom_point(aes(x = knots, y = AIC, col = "red")) +
  geom_point(aes(x = knots, y = BIC, col = "blue")) +
  ylab("BIC/AIC") +
  xlab("Number of b-spline internal knots") +
  scale_color_manual(labels = c("BIC", "AICc"), values = c("blue", "red")) +
  theme(legend.title = element_blank())
```


AIC suggests 2 knots for piecewise linear

### Altering degree of b-spline polynomial

Considering polynomials of degree 1 to 5

```{r, warning=FALSE, message=FALSE}
AIC_s <- rep(0,5)
BIC_s <- rep(0,5)

for (i in 1:5){
  
  PS_model <- lmer(data = fixation_intervals,
                       pupil_mean ~ bs(event_interval, degree = i, df = (i+2)) + (bs(event_interval, degree = i, df = (i+2)) | ID))

  AIC_s[i] <- AICc(PS_model)
  BIC_s[i] <- BIC(PS_model)
}

ggplot(data = data_frame(degree = 1:5, AIC = AIC_s, BIC = BIC_s)) +
  geom_point(aes(x = degree, y = AIC, col = "red")) +
  geom_point(aes(x = degree, y = BIC, col = "blue")) +
  ylab("BIC/AIC") +
  xlab("Degree of b-spline polynomial") +
  scale_color_manual(labels = c("BIC", "AIC"), values = c("blue", "red")) +
  theme(legend.title = element_blank())

AIC_s
```



AIC suggests a piece-wise linear model for the b-spline approximation.

